Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-10 paper/disc20clustering.tex
Comments
Self-review

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
\end{theorem}

@@

	\ALG performs at least one swap for every expansion of a ground set.
	Each expansion reduces the number of isolated nodes by $1$.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - l) / 2$.
\end{proof}

@@

\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial configuration $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure A final configuration $P_F$ 
		\STATE {For each node $v$ create a component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
		%\STATE{$P := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
		\IF{$C_1 \neq C_2$}
		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
		%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
		%		\COMMENT{i.e.~if not in the same cluster.}}       
		\IF{$C_1$ and $C_2$ are not collocated}
		\STATE {Move to a partition closest to $P_I$, respecting all components in $\mathcal{C}$.} 
		%		\COMMENT{move to a partition closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}

>>>>

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least $(k\cdot \ell - k - 2\ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
\end{theorem}

@@

	\ALG performs at least one swap for every expansion of a ground set.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on the initial cluster of $x_i$ is still a trivial component; this can occur at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
\end{proof}

@@

\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial configuration $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure A final configuration $P_F$ 
		\STATE {For each node $v$ create a component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
		%\STATE{$P := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
		\IF{$C_1 \neq C_2$}
		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
		%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
		%		\COMMENT{i.e.~if not in the same cluster.}}       
		\IF{$C_1$ and $C_2$ are not collocated}
		\STATE {Move to a partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
		%		\COMMENT{move to a partition closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}
